---
title: "Lab 06"
author: "Xiaofan Zhu"
date: "`r Sys.Date()`"
output: github_document 
always_allow_html: true
---

```{r install-libraries}
library(tidytext)
library(tidyverse)
library(dtplyr)
```
## Step 1. Read in the data
First download and then read in with read.csv()

```{r read-data, cache=TRUE}
if(!file.exists("mtsamples.csv")){
 download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv", 
               "mtsamples.csv", method="libcurl", timeout = 60)
}

mts <- read.csv("mtsamples.csv")
str(mts)
mts <- as_tibble(mts)
mts
```
## Question 1: What specialties do we have?
## We can use count() from dplyr to figure out how many records are there per specialty?
```{r}
specialties <- 
    mts %>%
    count(medical_specialty)

specialties %>%
    arrange(desc(n)) %>%
knitr::kable()
# There are 40 medical specialties.
```
```{r}
specialties %>%
  top_n(10) %>%
  ggplot(aes(x = n, y = fct_reorder(medical_specialty,n))) + 
   geom_col()
# The distribution is not at all uniform.
```
## Question 2 Visualize the top 20 most frequent words in the transcription column
## Tokenize the the words in the transcription column Count the number of times each token appears Visualize the top 20 most frequent words
```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  #anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

##Question 3. Redo analysis in Q2, and remove stopwords.

## Redo visualization but remove stopwords first. Bonus points if you remove numbers as well
```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  # use regular expression to filter out numbers
  filter( !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
# Removing the stopwords and numbers gives us a much better idea of what the text is about.
```
## Question 4
## repeat question 2, but this time tokenize into bi-grams.
```{r}
mts %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col()
```
```{r}
mts %>%
  unnest_ngrams(trigram, transcription, n=3) %>%
  count(trigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(trigram, n))) +
  geom_col()
# Top 20 trigrams seemed to return a few more medical word group than bigrams.
```

## Question 5 Using the results you got from questions 4. Pick a word and count the words that appears after and before it.

```{r}
ptbigram <-
  mts %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  separate(bigram,into = c("word1", "word2"), sep = " ") %>%
  select(word1, word2) %>%
  filter(word1 == "patient" | word2 == "patient")
# Words appearing before patient:
ptbigram %>%
  filter(word2=="patient") %>%
  count(word1, sort=TRUE) %>%
  anti_join(stop_words, by = c("word1" = "word")) %>%
  top_n(10) %>%
knitr::kable()

```

```{r}
# Find the words following patient:
ptbigram %>%
  filter(word1=="patient") %>%
  count(word2, sort=TRUE) %>%
  anti_join(stop_words, by = c("word2" = "word")) %>%
  top_n(10) %>%
knitr::kable()
# selecting by n


```

## Question 6 Which words are most used in each of the specialties. you can use group_by() and top_n() from dplyr to have the calculations be done within each specialty. Remember to remove stopwords. How about the most 5 used words?

```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  group_by(medical_specialty) %>% 
  count(word, sort = TRUE) %>%
  filter( !(word %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5, n) %>%
  arrange(medical_specialty, desc(n)) %>%
 knitr::kable()
```







