---
title: "assignment1"
date: "2022-09-23"
output: github_document 
always_allow_html: true
---
```{r}
library(data.table)
library(tidyverse)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(leaflet)
library(lubridate)
library(skimr)
library(readr)
```

## step 1 check the dimensions, headers, footers, variable names and variable types. Check for any data issues, particularly in the key variable we are analyzing.
```{r}
data_2004 <- data.table::fread("../assignment1/data_2004.csv")
data_2019 <- data.table::fread("../assignment1/data_2019.csv")

```
## check the dimensions and headers and footers of the data
```{r}
dim(data_2004)
head(data_2004)
tail(data_2004)

dim(data_2019)
head(data_2019)
tail(data_2019)

```
## Check the variable types in the data
```{r}
str(data_2004)
str(data_2019)
```

## exclude pm2.5<0
```{r}
data_2004 <- data_2004[data_2004$`Daily Mean PM2.5 Concentration`>=0,]
dim(data_2004)
```
```{r}
data_2019 <- data_2019[data_2019$`Daily Mean PM2.5 Concentration`>=0,]
dim(data_2019)
```

## summary data
```{r}
summary(data_2004)
summary(data_2004$daily_mean_pm2_5_concentration)
summary(data_2019)
summary(data_2019$daily_mean_pm2_5_concentration)
```
## check missing value
```{r}
sum(is.na(data_2004))
sum(is.na(data_2004$daily_mean_pm2_5_concentration))
sum(is.na(data_2019))
sum(is.na(data_2019$daily_mean_pm2_5_concentration))
# The number of rows in 2019 is greater than in 2004, which means that the 2019 dataset will carry more energy than 2004. When we examined the distribution of daily mean values of PM2.5 concentrations, we found that PM2.5 concentrations were negative. It turns out that if the atmosphere is clean, the concentration will be zero. But if there is noise in the measurement, there will be a negative number. So we need to get rid of these negative values. According to the daily mean distribution of PM2.5 concentration, the overall mean of PM2.5 concentration in 2019 was lower than that in 2004.
```



## Step 2 Combine the two years of data into one data frame. Use the Date variable to create a new column for year, which will serve as an identifier. Change the names of the key variables so that they are easier to refer to in your code.
```{r}
## creat a new column for year
data_2004[, year := 2004]
data_2019[, year := 2019]
pm_data <- rbind(data_2004, data_2019)

```

## Change the names

```{r}
pm_data <- as_tibble(pm_data)
```

```{r}
pm_data <- pm_data %>%
  rename(pm25 = `Daily Mean PM2.5 Concentration`,
         countycode = COUNTY_CODE,
         county = COUNTY,
         site = `Site Name`,
         lat = SITE_LATITUDE,
         lon = SITE_LONGITUDE,
         )
```
## Step 3 Create a basic map in leaflet() that shows the locations of the sites (make sure to use different colors for each year). Summarize the spatial distribution of the monitoring sites.
```{r}
temp.pal <- colorFactor(c("red", "blue"), domain = pm_data$year)
leaflet(pm_data) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircles(
    lat = ~lat, lng = ~lon,
    label = ~year, color = ~temp.pal(year),
    opacity = 1, fillOpacity = 1,radius = 500  ) %>%
   addLegend('bottomleft', pal = temp.pal, values = pm_data$year,
            title = 'year', opacity = 1)
# Looking at the site location map, there are more sites in the 2019 dataset than in 2004. In both years, observation sites were sparsely distributed. Most of the sites are on the coast, and these are usually developed cities with lots of people living there. Due to the influence of human activities, PM2.5 concentrations may be higher than in untouched areas.
```

## step 4 Check for any missing or implausible values of PM 2.5 in the combined dataset. Explore the proportions of each and provide a summary of any temporal patterns you see in these observations.
```{r}
sum(is.na(pm_data$pm25))

summary(pm_data$pm25)
# When we examined the distribution of daily mean values of PM2.5 concentrations, we found that PM2.5 concentrations were negative. It turns out that if the atmosphere is clean, the concentration will be zero. But if there is noise in the measurement, there will be a negative number. So we need to get rid of these negative values. In terms of the proportion of PM2.5 concentration, the highest concentrations were 4.4 to 11.3(first to third quartiles). Areas with extremely high PM2.5 concentrations may be biased, leading to higher average concentration levels.
```
## step 5 Explore the main question of interest at three different spatial levels.

```{r}
# state
library(ggplot2)
ggplot(data = pm_data)+
  geom_boxplot(mapping=aes(x=year, y=log10(pm25), fill=year,group =year))
```






